# Installation instructions

Note that these instructions will be updated based on participants'
feedback.

**Last update**: `r date()`

## Software and hardware requirement

**Participants are required** to bring their own laptop with
[R](https://www.r-project.org/) version 3.6.0 or greater. We also
recommend to also install the latest version of
[RStudio](https://www.rstudio.com/products/rstudio/download/).

Please make sure that your computer’s hardware is sufficiently powered
(>4 GB RAM, > 2 GB free disk space), that you have a working wireless
card, and that you have administrator rights.

## Installation script

To install all requirements, please copy and paste this line of code
in your R console.

```{r installscript, eval = FALSE, purl = FALSE}
source("https://uclouvain-cbio.github.io/BSS2019/installation.R")
```

The whole installation can take some time, if many of the required
packages aren't already installed. At the end, you should see a
*Success!* message, indicating that all dependencies are installed and
up-to-date, and you will also have the `airway.zip` and `extdata.zip`
files in your working directory.

The rest of this page describes the content of the installation script
executed with the above line. It is given for information. In case of
installation issues, check you internet connection and try to run the
script a second time. If problems persist, please send your questions
to the summer school mailing list, or try to execute the installation
steps below one by one.

## R version

```{r rversion, eval = TRUE}
message("* Checking R version.")
stopifnot(base::version$major == "3")
stopifnot(base::version$minor >= "6.2")
```

## Package installation

We are going to user the `BiocManager` package to install all other
packages from the `CRAN` and `Bioconductor` repositories. Let's start
by installing `BiocManager` if it's not already available:

```{r biocmanagermsg, eval = TRUE}
message("* Installing BiocManager.")
```

```{r biocmanager, eval = TRUE}
if (!require("BiocManager"))
    install.packages("BiocManager")
```

Below is a vector containing the names of all required packages:


```{r pkgs}
pkgs <- c("rjson", "shinydashboard", "shiny", "shinyjs",
          "shinythemes", "shinyFiles", "openxlsx", "statOmics/MSqRob",
          "iSEE", "SummarizedExperiment", "SingleCellExperiment",
          "tximeta", "org.Hs.eg.db", "DESeq2", "edgeR", "pheatmap",
          "rmdformats", "apeglm", "lme4", "tidyverse", "rmarkdown",
          "Biobase", "DT", "fdrtool", "graphics", "grDevices",
          "limma", "MASS", "Matrix", "methods", "MSnbase", "numDeriv",
          "openxlsx", "plyr", "preprocessCore", "statmod", "msdata",
          "svDialogs", "utils", "vsn", "miniUI", "htmltools", "mzR",
          "ProtGenerics", "BiocParallel", "snow", "httpuv", "xtable",
          "htmlwidgets", "colorspace", "RColorBrewer", "dichromat",
          "munsell", "labeling", "affyio", "zlibbioc", "doParallel",
          "gtable", "scales", "tibble", "lazyeval", "reshape2",
          "affy", "impute", "pcaMethods", "mzID", "MALDIquant",
          "ggplot2", "XML", "minqa", "nloptr", "magrittr", "yaml",
          "stringr", "stringi", "foreach", "iterators", "Rcpp",
          "BiocGenerics", "BiocManager", "biomaRt", "Biostrings",
          "BSgenome", "BSgenome.Hsapiens.UCSC.hg38", "ggbeeswarm",
          "ISLR", "ca", "class", "curatedTCGAData", "dplyr",
          "EnrichmentBrowser", "GenomicAlignments", "GenomicFeatures",
          "GenomicRanges", "ggplot2", "GO.db", "IRanges", "KEGGREST",
          "MultiAssayExperiment", "org.Hs.eg.db", "rtracklayer",
          "limma", "SingleCellExperiment", "SummarizedExperiment",
          "tibble", "tidyr", "TxDb.Hsapiens.UCSC.hg38.knownGene")
pkgs <- sort(unique(pkgs))
pkgs
```

We first check which packages are already available, update all old
packages with the latest available versions and proceed with the
installation of packages that aren't installed yet.

```{r pkgsmsg, eval = TRUE}
message("* Installing packages.")
```


```{r install, eval = TRUE}
ip <- rownames(installed.packages())
pkgs_to_install <- pkgs[which(!sub("^.+/", "", pkgs) %in% ip)]
BiocManager::install(pkgs_to_install, update = TRUE, ask = FALSE, upgrade = 'always')
```

<!-- Below, we also install the `BiocIntro` package if not already -->
<!-- available: -->

<!-- ```{r biocintro, eval = TRUE} -->
<!-- if (!"BiocIntro" %in% ip) -->
<!--     BiocManager::install("Bioconductor/BiocIntro", ref = "BSS2020") -->
<!-- ``` -->

## Download additional data

The
[`airway.zip`](https://uclouvain-cbio.github.io/BSS2019/airway.zip)
file contains data that will be used during the RNASeq sessions. It can
be downloaded from R in the working directory with:

```{r datamsg, eval = TRUE}
message("* Downloading data.")
```

```{r airway, eval = TRUE}
download.file("https://uclouvain-cbio.github.io/BSS2019/airway.zip", "airway.zip")
download.file("https://uclouvain-cbio.github.io/BSS2019/extdata.zip", "extdata.zip")
```

## Success

If you see the following message in your console, it means you have
successfully installed all requirements.

```{r success, eval = TRUE}
message("* Success!")
```

Back to the [summer school page](./index.html).
